\Title

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : C

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. This results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$: 
%
\begin{linenomath*}
\begin{equation}\label{eq:stream_b}
\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
\end{equation}
\end{linenomath*}
%
In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons. In simulations, however, it should be properly reordered. The activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions (\seeFig{izhikevich}). Crucially, when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, the discharge probability increases, and especially when these spikes converge at the soma in a synchronous manner. We have defined the HD-SNN model in all generality in this subsection, and in the next subsection we describe an implementation of our model adapted to the motion detection task. 
%
\subsection{Application of the HD-SNN model to the motion detection task}
%
It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision. In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}. To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.

Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented. In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions. To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$. The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$. Using this dense representation, the processing of $A$ by the model can be written as: %
%
\begin{linenomath*}
\begin{equation}\label{eq:kernel_b}
\forall x, y, \timev, \; \;
\mathcal{C}_\class(x, y, \timev)
= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
\end{equation}
\end{linenomath*}
%
where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time. This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream. A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task. By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.

>>>>

Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$. As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field. This results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$: 
%
\begin{linenomath*}
\begin{equation}\label{eq:stream_b}
\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
\end{equation}
\end{linenomath*}
%
In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons. In simulations, however, it should be properly reordered. The activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions (\seeFig{izhikevich}). Crucially, when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, the discharge probability increases, and especially when these spikes converge at the soma in a synchronous manner. We have defined the HD-SNN model in all generality in this subsection, and in the next subsection we describe an implementation of our model adapted to the motion detection task. 
%
\subsection{Application of the HD-SNN model to the motion detection task}
%
It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision. In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$. The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$. First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}. This takes advantage of the position invariance observed in images and exploited in convolutional neural networks, and we can further assume that synaptic motifs should be similar across different positions. As a consequence, we can define a spatio-temporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.

Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ such that all possible delays associated to the different presynaptic addresses are represented. In particular, for each class $\class$ of the supervision task, the whole synaptic set can be represented as one kernel which is represented by the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions. To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$. By convenienece, the connectivity of neuron $\postsynaddr$ is defined locally, around its position $(x_\postsynaddr, y_\postsynaddr)$. Using this dense representation, the processing of $A$ by the model can be written as: %
%
\begin{linenomath*}
\begin{equation}\label{eq:kernel_b}
\forall x, y, \timev, \; \;
\mathcal{C}_\class(x, y, \timev)
= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
\end{equation}
\end{linenomath*}
%
where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration). Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time. This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream. A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}. Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task. By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
